% Midterm.
% Humam Rashid
% Spring 2020, CISC 7214X, Prof. Cox

\documentclass{scrartcl}
\usepackage{amsmath, amsthm, amssymb}
\usepackage[a4paper, margin=1in]{geometry}
\usepackage[shortlabels]{enumitem}
\begin{document}
\begin{flushleft}
    Humam Rashid\\
    Midterm, Spring 2020\\
    CISC 7214X, Prof. Cox
\end{flushleft}
    \bigskip
    
    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    % Problem 1 on solving summations.
    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    \noindent
    1. Solve the summation.
    \begin{align*}
        \sum_{i=1}^{n} 4i
    \end{align*}
    \textit{Solution}:
    \begin{align*}
        \sum_{i=1}^{n} 4i &= 4 \sum_{i=1}^{n} i\\[6pt]&\text{by distributive identity}\\[6pt]
                          &= 4 \bigg(\frac{n (n + 1)}{2}\bigg)\\[6pt]&\text{by sum of first $n$
                          natural numbers}\\[6pt]
                          &= 2n(n + 1)
    \end{align*}

    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    % Problem 2 on solving and proving recurrence equations.
    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    \bigskip
    \noindent
    2. Solve the following recurrence and prove the solution correct with induction:
    \begin{align*}
        T(1) &= 0\\
        T(n) &= n + 2 T\bigg(\frac{n}{2}\bigg)
    \end{align*}
    \textit{Solution}: $T(n) = n\;lg(n)$, claimed. We see that $T(1) = 1\;lg(1) = 0$.\\
    \begin{proof}
        For $k < n$, we have:
        \begin{align*}
            T(k) &= k\;lg(k)\\
            T(n) &= n + 2 T\bigg(\frac{n}{2}\bigg)\\
                 &= n + 2 \bigg(\frac{n}{2}\;lg(n) - \frac{n}{2}\bigg)\\
                 &= n + n\;lg(n) - n\\
                 &= n\;lg(n)
        \end{align*}
    \end{proof}
    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    % Problem 3 on data structure for insertion and max extraction for set of integers.
    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    \newpage
    \noindent
    3. Suppose you want to be able to perform the following 2 operations on a set of integers.
    Insert a new item and remove the maximum item. Describe a data structure that supports this and
    analyze its running time.\\

    \noindent
    \textit{Answer}: One data structure that readily supports the given requirements is a
    \emph{Max-Priority Queue}. A Max-Priority Queue can be built on top of a Max-Heap. Building the
    Max-Heap from a set of $n$ unsorted integers takes $O(n)$ time. Removing the maximum item is an
    $O(lg\,n)$ operation since returning the maximum can be done in constant time but maintaining
    the Max-Heap takes $O(lg\,n)$ time. Inserting an item into a Max-Heap with $n$ items also takes
    $O(lg\,n)$ because of the Max-Heap maintenance procedure. (I am using Max-Priority Queue here as
    defined in Cormen et al., effectively making it a simple abstraction over a Max-Heap).\\

    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    % Problem 4 on algorithm for finding shortest-path tree in a network.
    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    \bigskip
    \noindent
    4. Suppose you are given a network with $n$ cities, and a direct connection between each pair of
    cities with the distance between them. Given a starting city $s$, describe an algorithm to find
    the shortest path from $s$ to each city in the network. Analyze the complexity.\\

    \noindent
    \textit{Answer}: An effective algorithm to use here is \emph{Dijkstra's Shortest Path
    Algorithm}. This algorithm can be used to construct a shortest-path tree from the network of $n$
    cities, using $s$ as the starting node. The time complexity of Dijkstra's shortest path
    algorithm is dependent on the data structure used to implement the procedure that gives us the
    shortest-path estimate for a vertex $u \in V - S$, where $V$ is the set of vertices representing
    the cities (i.e., $|V| = n$) and $S$ is the set of vertices for which the final shortest-path
    weights from $s$ are already determined. Given a procedure \verb|EXTRACT-MIN()| to get $u$, the
    following describes several well known methods of implementing the procedure and the resulting
    overall time complexity of Dijkstra's shortest path algorithm ($E$ is the set of edges in the
    network):
    \begin{enumerate}[(a)]
        \item If \verb|EXTRACT-MIN| is implemented using a simple linked-list or array requiring
            linear search to find the minimum, Dijkstra's shortest path algorithm has time
            complexity: $O(|V|^2)$ or $O(n^2)$.\\
        \item If \verb|EXTRACT-MIN| is implemented using a Binary-Min Heap, then Dijkstra's shortest
            path algorithm has time complexity: $O((|V| + |E|)\;lg\,|V|)$ or $O((n + |E|)\;lg\,n)$.
            For connected networks (as in our case), this further reduces to $O(|E|\;lg\,|V|)$ or
            $O(|E|\;lg\,n)$.\\
        \item If \verb|EXTRACT-MIN| is implemented using a Fibonacci Heap, then Dijkstra's shortest
            path algorithm has time complexity: $O(|V|\;lg\,|V| + |E|)$ or $O(n\;lg\,n + |E|)$.
    \end{enumerate}

    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    % Problem 5 on complexity classes.
    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    \bigskip
    \noindent
    5. Given a graph $G$ as input, \textsf{stCONN} is the problem of determining if there is a path
    from start vertex $s$ to target vertex $t$. Sketch a proof that \textsf{stCONN} is in the class
    \textsf{NL}.\\

    \noindent
    \textit{Answer}: A nondeterministic log space Turing machine (NLTM) deciding \textsf{stCONN}
    starts at $s$ and nondeterministically guesses the vertices of a path from $s$ to $t$. The NLTM
    only records the position of the current vertex at each step on the work tape, not the entire
    path (so that it is restricted to logarithmic space usage). The NLTM nondeterministically
    selects the next vertex from those vertices the current vertex directly points to. This action
    is repeated until it reaches $t$ and the NLTM accepts or until it has gone for $m$ steps and
    rejects, where $m$ is the number of vertices in the graph. This process shows that
    \textsf{stCONN} is in \textsf{NL}.

    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    % Problem 6 on complexity classes.
    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    \bigskip
    \noindent
    6. Show that \textsf{NL} $\subseteq \mathcal{P}$.\\

    \noindent
    \textit{Answer}:

    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    % Problem 7 on complexity classes.
    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    \newpage
    \noindent
    7. How do we know that \textsf{NPSPACE $=$ PSPACE}?\\

    \noindent
    \textit{Answer}: This is a direct result of Savitch's theorem, which says \textsf{NSPACE$(s(n))
    \subseteq$ SPACE$(s(n)^2)$}. That is, for space complexity, any nondeterministic Turing machine
    that uses $s(n)$ space can be converted to a deterministic Turing machine that uses only
    $s(n)^2$ space. \textsf{NPSPACE $=$ PSPACE} is a direct result based on the fact that the square
    of any polynomial is still a polynomial.\\

    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    % Problem 8 on complexity classes.
    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    \bigskip
    \noindent
    8. Let $X$ be a problem that is in the class $\mathcal{NP}$. For each of the following, write if
    it is ``True,'' ``False,'' or ``We don't know.'' Justify your answer.
    \begin{enumerate}[(a)]
        \item Solutions to $X$ can be verified with a polynomial-time algorithm.\\

            \textit{True}. This is directly from the definition of $\mathcal{NP}$.\\
        \item There is no polynomial-time algorithm for $X$.\\

            \textit{We don't know}. If it was known that $X \in \mathcal{P} \subseteq \mathcal{NP}$,
            then the claim would have been false (i.e., it would give us confidence that such an
            algorithm exists), but for the more general case of $X \in \mathcal{NP}$, the answer is
            still uncertain (until and unless it is also proven that $\mathcal{NP} \subseteq
            \mathcal{P} \Rightarrow \mathcal{P} = \mathcal{NP}$).\\
        \item If $X$ is $\mathcal{NP}$-Hard, then $X$ is $\mathcal{NP}$-Complete.\\

            \textit{True}. An $\mathcal{NP}$-Complete problem is one that is $\mathcal{NP}$-Hard and
            also in $\mathcal{NP}$, therefore this is directly from the definition of
            $\mathcal{NP}$-Complete.\\
        \item If there exists a polynomial-time algorithm for \textsf{SAT}, then there exists a
            polynomial-time algorithm for $X$.\\

            \textit{True}. The \textsf{SAT} problem is known to be $\mathcal{NP}$-Complete. Since
            $\mathcal{NP}$-Completeness implies the problem is both in $\mathcal{NP}$ and
            $\mathcal{NP}$-Hard, all problems in $\mathcal{NP}$ must be polynomial-time Karp
            reducible to \textsf{SAT} (by definition of $\mathcal{NP}$-Hard). In other words,
            \textsf{SAT} is ``at least as hard'' as $X$ and therefore \textsf{SAT} $\in \mathcal{P}
            \Rightarrow X \in \mathcal{P}$.
    \end{enumerate}

\end{document}

% EOF.
